---
title: "Linkedin postings"
subtitle: "Part one"
author: "Sean Ng"
organization: "AIMdata"
date-modified: "11 January 2026"
execute: 
  echo: false
---


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 9)


library(tidyverse)
library(here)
library(janitor)
library(scales)
library(tidytext)
library(widyr)
library(ggraph)
library(patchwork)
library(kableExtra)
library(viridis)
library(DT)
library(skimr)
library(jsonlite)
library(jsonld)

`%out%` <- Negate(`%in%`)
options(scipen = 100)
theme_set(theme_light())
range_wna <- function(x){(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))}

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```

```{r data}

industries <- read_csv("./data/companies/company_industries.csv") |> 
  mutate(industry_simple = 
           case_when(
             industry %in% c("Staffing and Recruiting", 
                             "Human Resources Services") ~ 
               "Recruiting", 
             industry %in% c("Farming", "Fisheries") ~ "Food Production", 
             industry %in% c("Business Consulting and Services", 
                             "Research Services", 
                             "Professional Training and Coaching", 
                             "Outsourcing and Offshoring Consulting",
                             "Government Relations Services", 
                             "Market Research", 
                             "Strategic Management Services", 
                             "Executive Offices", 
                             "Design Services") ~ 
               "Business Consulting", 
             industry %in% c("Public Relations and Communications Services", 
                             "Events Services", 
                             "Printing Services", 
                             "Graphic Design") ~ 
               "Communications and Events", 
             industry %in% c("IT Services and IT Consulting", 
                             "Software Development", 
                             "Computer and Network Security", 
                             "Technology, Information and Internet", 
                             "Information Services", 
                             "Computer Games", 
                             "Computer Networking Products", 
                             "Mobile Gaming Apps") ~ 
               "IT", 
             industry %in% c("Financial Services", 
                             "Banking", 
                             "Venture Capital and Private Equity Principals", 
                             "Investment Banking", 
                             "Investment Management", 
                             "Accounting") ~ 
               "Financial Services",  
             industry %in% c("Construction", 
                             "Wholesale Building Materials", 
                             "Civil Engineering", 
                             "Architecture and Planning") ~ 
               "Construction", 
             industry == "Advertising Services" ~ "Advertising", 
             industry %in% c("Non-profit Organizations", 
                            "Civic and Social Organizations", 
                            "Religious Institutions", 
                            "International Trade and Development", 
                            "Political Organizations", 
                            "Fundraising", 
                            "International Affairs", 
                            "Think Tanks", 
                            "Philanthropic Fundraising Services", 
                            "Public Policy Offices") ~ 
               "Civil Society", 
             industry %in% c("Real Estate", 
                             "Facilities Services", 
                             "Environmental Services", 
                             "Recreational Facilities", 
                             "Leasing Non-residential Real Estate") ~ 
               "Real Estate", 
             industry == "Insurance" ~ "Insurance", 
             industry %in% c("Law Practice", 
                             "Legal Services", 
                             "Administration of Justice", 
                             "Alternative Dispute Resolution") ~ 
               "Legal", 
             industry %in%  c("Higher Education", 
                              "Education Administration Programs", 
                              "Primary and Secondary Education", 
                              "E-Learning Providers") ~ 
               "Education", 
             industry %in% c("Government Administration", 
                             "Legislative Offices", 
                             "Individual and Family Services") ~ 
               "Government Administration",
             industry %in% c("Biotechnology Research", 
                             "Pharmaceuticals Manufacturing", 
                             "Nanotechnology Research") ~ 
               "Pharmaceuticals and Biotech", 
             industry %in% c("Food and Beverage Services", 
                             "Restaurants", 
                             "Beverage Manufacturing", 
                             "Dairy Product Manufacturing") ~ 
               "F&B", 
             industry %in% c("Mental Health Care",  
                             "Hospitals and Health Care", 
                             "Medical Practices")
             ~ "Healthcare", 
             industry %in% c("Medical Equipment Manufacturing", 
                             "Industrial Machinery Manufacturing", 
                             "Appliances, Electrical, and Electronics Manufacturing", 
                             "Food and Beverage Manufacturing",
                             "Pharmaceutical Manufacturing", 
                             "Defense and Space Manufacturing", 
                             "Chemical Manufacturing", 
                             "Motor Vehicle Manufacturing", 
                             "Renewable Energy Semiconductor Manufacturing", 
                             "Aviation and Aerospace Component Manufacturing", 
                             "Automation Machinery Manufacturing", 
                             "Packaging and Containers Manufacturing", 
                             "Semiconductor Manufacturing", 
                             "Personal Care Product Manufacturing", 
                             "Computers and Electronics Manufacturing", 
                             "Plastics Manufacturing", 
                             "Furniture and Home Furnishings Manufacturing", 
                             "Paper and Forest Product Manufacturing", 
                             "Textile Manufacturing",
                             "Computer Hardware Manufacturing", 
                             "Sporting Goods Manufacturing", 
                             "Glass, Ceramics and Concrete Manufacturing", 
                             "Railroad Equipment Manufacturing", 
                             "Shipbuilding", 
                             "Tobacco Manufacturing", 
                             "Machinery Manufacturing", 
                             "Manufacturing")  ~ 
               "Manufacturing", 
             industry %in% c("Oil and Gas", 
                             "Mining") ~ 
               "Natural Resources", 
             industry %in% c("Utilities", 
                             "Telecommunications", 
                             "Wireless Services") ~ 
               "Telecoms and Utilities", 
             industry %in% c("Entertainment Providers", 
                             "Spectator Sports", 
                             "Broadcast Media Production and Distribution", 
                             "Media Production", 
                             "Book and Periodical Publishing",
                             "Gambling Facilities and Casinos", 
                             "Museums, Historical Sites, and Zoos", 
                             "Online Audio and Video Media", 
                             "Musicians", 
                             "Performing Arts", 
                             "Translation and Localization", 
                             "Newspaper Publishing", 
                             "Artists and Writers", 
                             "Libraries", 
                             "Photography",
                             "Animation and Post-production", 
                             "Writing and Editing") ~ 
               "Entertainment", 
             # Tell me it's not true
             industry %in% c("Airlines and Aviation", 
                             "Transportation, Logistics, Supply Chain and Storage", 
                             "Freight and Package Transportation", 
                             "Maritime Transportation", 
                             "Warehousing and Storage", 
                             "Truck Transportation") ~ 
               "Transportation and Logistics", 
             industry %in% c("Retail", 
                             "Retail Apparel and Fashion", 
                             "Consumer Services", 
                             "Retail Luxury Goods and Jewelry", 
                             # Maybe contentious, but Costco is a retailer
                             "Wholesale", 
                             "Retail Office Equipment",
                             "Retail Groceries", 
                             "Wholesale Import and Export") ~ 
               "Retail and Sales", 
             industry %in% c("Security and Investigations", 
                             "Public Safety", 
                             "Law Enforcement", 
                             "Armed Forces") ~ "Security", 
             industry %in% c("Travel Arrangements", 
                             "Hospitality") ~ 
               "Travel",
             industry %in% c("Veterinary Services", 
                             "Alternative Medicine", 
                             "Wellness and Fitness Services") ~ "Wellness"
           )) |> 
  # Just dropping all the extras
  group_by(company_id) |> 
  slice_head(n = 1) |> 
  ungroup()

job_category_cleaning <- function(tbl){
  tbl |>  
    mutate(
      job_category = 
        case_when(
          str_detect(title, "Trainee|Intern|Paralegal|Entry Level|Support|Teller|ENTRY LEVEL|Accounting Clerk|Graduate Program|Universal Banker|Data Entry Clerk") ~ 
                    "Entry Level",
          str_detect(title,
                     "Manager|MANAGER|Superintendent|Supervisor|Lead|Product Owner|Scrum Master|Cloud Architect|Sous Chef")  & !str_detect(title, "Chief")
                  ~ "Manager", 
                  str_detect(title, "Officer|Social Worker|Teacher|TEACHER|teacher") & !str_detect(title, "Chief") ~ 
                    "Officer", 
                  str_detect(title, "Engineer|Developer|Quality Inspector|Quality Control Inspector|Construction Inspector") ~ 
                    "Engineer", 
                  str_detect(title, "Technician|Detailer|Mechanic|Forklift Operator|Electrician|Machinist|Equipment Operator|Carpenter|Assembler|Operator|Medical Biller|Hair Stylist|Coordinator|Dispatcher| Tech") ~ 
                    "Technician", 
                  str_detect(
                    title, 
                    "Specialist|Technologist|Controller|Phlebotomist|Scientist|Pathologist|Estimator|Data Architect|System Administrator|Systems Administrator|Database Administrator|Planner|Java   Architect|Network Administrator|Salesforce Administrator") ~ 
                    "Specialist", 
                  str_detect(title, "Analyst|Data Modeler") ~ "Analyst",
                  str_detect(title, "Cook|Driver|Housekeeper|Cashier|Dishwasher|Server|Bartender|Warehouse Worker|Janitor|Groomer|Food Service Worker|Welder|Custodian|Laborer|Dashers|Housekeeping   Aide|Barista|In-Store Shopper|Handler|Stocker|Drive with DoorDash|Caregiver|Warehouse Part Time Overnight|Cleaner|Warehouse Part Time Days|Restaurant Delivery|Warehouse Order Selector|Groundskeeper|Keyholder|Porter|Valet|store driver|Concierge|Room Attendant") ~ 
                    "Labourer", 
                  str_detect(title, "Consultant|Advisor") ~ 
                    "Consultant", 
                  str_detect(title, "Chief|Director|Business Owner|Co-Founder|Executive|Vice President|Head") &
            ! str_detect(title, "Sales") ~ 
                    "Director",
                  str_detect(title, 
                             "Assistant|Receptionist|Retail Sales – Part Time|Front Desk Agent|ASSISTANT|Aide|Office Administrator|Contract Administrator|Legal Secretary|Branch Office Administrator|Payroll Administrator") ~ "Assistant",
                  # Not sure about "Buyer"
                  str_detect(
                    title, "Associate|Salesperson|Team Member|Generalist|Merchandiser|Human Resources   Business Partner|Buyer|Relationship Banker|salesperson|Retail Parts Pro|ASSOCIATE|Sales Agent|Sales Professional|Sleep Expert - Sales|Representative|REPRESENTATIVE| Rep|Personal Trainer|PROMOTIONS AND SALES POSITIONS|Insurance Agent|Personal Banker|Recruiter") ~ 
                    "Associate",
                  str_detect(title, 
                             "Attorney|Nurse|Therapist|Dental Hygienist|Architect|Licensed Clinical Social Worker|Physician|Nurse|NURSE|RN|Registered Dietitian|Chemist|Dentist|OCCUPATIONAL THERAPIST|Licensed Mental Health Counselor|Pharmacist|Psychiatrist|Veterinarian|Accountant|Bookkeeper|Auditor") ~ 
                    "Licensed Professional",
                  str_detect(title, "Writer|writer|Content Editor|Designer") ~ "Creative",
                  TRUE ~ "Other"
                  )) |> 
    mutate(senior = ifelse(str_detect(title, "Senior"), 
                           "senior", 
                           "not senior")) 
}


postings_clean <- read_csv("./raw_data/postings.csv") |>
  filter(!is.na(company_id)) |> 
  left_join(
    industries,
    by = "company_id"
  ) |> 
  filter(!is.na(industry)) |> 
  filter(!is.na(company_name) & company_name != "Confidential") |> 
  # Likely typos since there's entry-level and associates
  # Fine to drop them since it's less than 10 entries
  filter(normalized_salary < 1000000 | is.na(normalized_salary)) |> 
  job_category_cleaning()
```

I think let's ignore all staffing and recruiting postings. 



```{r}
postings_clean |> 
  filter(industry_simple == "IT") |> 
  group_by(industry) |> 
  summarise(postings = n_distinct(job_id), 
            companies = n_distinct(company_id)) |> 
  ggplot(aes(x = postings, 
             y = fct_reorder(industry, postings))) +
  geom_col(aes(fill = companies))

```
## By industry

```{r}
postings_clean |> 
  filter(industry_simple != "Recruiting") |> 
  group_by(industry_simple) |> 
  summarise(postings = n_distinct(job_id), 
            companies = n_distinct(company_id)) |> 
  ggplot(aes(x = postings, 
             y = fct_reorder(industry_simple, postings))) +
  geom_col(aes(fill = companies)) + 
  scale_x_continuous(labels = comma) +
  scale_fill_viridis(direction = -1, option = "cividis") +
  labs(title = "LinkedIn job postings in 2023-2024, by industry", 
       subtitle = "Excludes postings by recruiters and middlemen", 
       y = "")
```

Manager, Officer, Associate, Engineer, Coordinator, Representative, Accountant, Assistant, Executive

Payscale by job category?


We can at least tell that our 

```{r}
postings_clean |> 
  filter(normalized_salary > 1000 | is.na(normalized_salary)) |> 
  group_by(job_category) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE), 
            min_salary = min(normalized_salary, na.rm = TRUE), 
            max_salary = max(normalized_salary, na.rm = TRUE)) |> 
  arrange(desc(salary))
```


```{r}
postings_clean |>
  # Not sure this is the right filter
  # We can play around with it
  filter(normalized_salary > 1000 | is.na(normalized_salary)) |> 
  group_by(job_category) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE)) |> 
  ggplot(aes(x = n, 
             y = fct_reorder(job_category, n))) + 
  geom_col(aes(fill = salary)) +
  geom_text(aes(label = comma(salary)), 
            hjust = "inward") +
  scale_fill_viridis(direction = -1, labels = comma) + 
  scale_x_continuous(labels = comma) + 
  labs(x = "Number of postings", 
       y = "", 
       title = "Mean salary by job category", 
       fill = "Mean\nSalary")

```

```{r}
postings_clean |>
  # Not sure this is the right filter
  # We can play around with it
  filter(normalized_salary > 1000 | is.na(normalized_salary)) |> 
  group_by(job_category) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE)) |> 
  arrange(desc(salary))
```

<br>

```{r}
postings_clean |> 
  filter(industry_simple != "Recruiting") |> 
  group_by(job_category, industry_simple) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE)) |> 
  group_by(industry_simple) |> 
  mutate(pc = n / sum(n),
         sort = sum(n[job_category %in% c(
           "Director", "Engineer", "Licensed Professional", "Manager")]) / 
           sum(n)) |> 
  ungroup() |> 
  mutate(job_category = fct_relevel(
    job_category, c("Director", "Engineer", "Licensed Professional", 
                    "Manager", "Consultant", "Creative", 
                    "Analyst", "Other", "Specialist", 
                    "Officer", "Associate", "Entry Level", 
                    "Technician", "Assistant", "Labourer")
  )) |> 
  arrange(desc(job_category)) |> 
  ggplot(aes(x = pc, y = fct_reorder(industry_simple, sort), 
             group = job_category)) + 
  geom_col(aes(fill = job_category)) + 
  scale_fill_viridis_d(option = "turbo") + 
  scale_x_continuous(labels = percent) +
  labs(title = "Breakdown of job postings by industry", 
       subtitle = "Sorted by percentage of 'good jobs'", 
       x = "", y = "", 
       fill = "") + 
  theme(legend.text = element_text(size = 9))
```

<br>

Education is dying (perhaps as a backlash towards its intensive monetisation and the purposeful destruction of the teaching profession), so is the legal industry, as well as government administration, which I can attest to have been in a state of atrophy even before DOGE and the great sack of the American government. These industries have an even worse proportion of good jobs than the oft-memed about Travel industry. 

We should probably cross-reference this with data scraped from Indeed, but the only available dataset is only a marketing sample of 30,000 jobs. 

<br>


```{r}
postings_clean |> 
  filter(industry_simple == "Communications and Events") |> 
  group_by(industry, job_category) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE))
```





```{r}
postings_clean |>
  job_category_cleaning() |> 
  filter(job_category == "Manager") |> 
  count(title, sort = TRUE)
  filter(is.na(job_category)) |> 
  arrange(desc(n))
```

```{r}
postings_clean |> 
  job_category_cleaning() |> 
  filter(normalized_salary < 500000 ) |> 
  ggplot(aes(x = normalized_salary)) + 
  geom_histogram() 
  select(title, max_salary, med_salary, min_salary, pay_period, normalized_salary)
```

<br>

53% of LinkedIn users earn more than USD 100,000 a year. 54% of college graduates use LinkedIn. 

https://blog.hootsuite.com/linkedin-demographics/#:~:text=What%20the%20data%20tells%20us,makers%20and%20executive%20audiences%20worldwide.


## Indeed

```{r}
indeed <- stream_in(file(("./data/marketing_sample_for_indeed_usa-indeed_usa_job__20210101_20210331__30k_data.ldjson")))

indeed |> glimpse()

indeed |> count(category, sort = TRUE)
```


