---
title: "Linkedin postings"
subtitle: "Part one"
author: "Sean Ng"
organization: "AIMdata"
date-modified: "11 January 2026"
execute: 
  echo: false
---


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 9)


library(tidyverse)
library(here)
library(janitor)
library(scales)
library(tidytext)
library(widyr)
library(ggraph)
library(patchwork)
library(kableExtra)
library(viridis)
library(DT)
library(skimr)
library(jsonlite)
library(jsonld)

`%out%` <- Negate(`%in%`)
options(scipen = 100)
theme_set(theme_light())
range_wna <- function(x){(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))}

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```

```{r data}

industries <- read_csv("./data/companies/company_industries.csv") |> 
  mutate(industry_simple = 
           case_when(
             industry %in% c("Staffing and Recruiting", 
                             "Human Resources Services") ~ 
               "Recruiting", 
             industry %in% c("Farming", "Fisheries") ~ "Food Production", 
             industry %in% c("Business Consulting and Services", 
                             "Research Services", 
                             "Professional Training and Coaching", 
                             "Outsourcing and Offshoring Consulting",
                             "Government Relations Services", 
                             "Market Research", 
                             "Strategic Management Services", 
                             "Executive Offices", 
                             "Design Services") ~ 
               "Business Consulting", 
             industry %in% c("Public Relations and Communications Services", 
                             "Events Services", 
                             "Printing Services", 
                             "Graphic Design") ~ 
               "Communications and Events", 
             industry %in% c("IT Services and IT Consulting", 
                             "Software Development", 
                             "Computer and Network Security", 
                             "Technology, Information and Internet", 
                             "Information Services", 
                             "Computer Games", 
                             "Computer Networking Products", 
                             "Mobile Gaming Apps") ~ 
               "IT", 
             industry %in% c("Financial Services", 
                             "Banking", 
                             "Venture Capital and Private Equity Principals", 
                             "Investment Banking", 
                             "Investment Management", 
                             "Accounting") ~ 
               "Financial Services",  
             industry %in% c("Construction", 
                             "Wholesale Building Materials", 
                             "Civil Engineering", 
                             "Architecture and Planning") ~ 
               "Construction", 
             industry == "Advertising Services" ~ "Advertising", 
             industry %in% c("Non-profit Organizations", 
                            "Civic and Social Organizations", 
                            "Religious Institutions", 
                            "International Trade and Development", 
                            "Political Organizations", 
                            "Fundraising", 
                            "International Affairs", 
                            "Think Tanks", 
                            "Philanthropic Fundraising Services", 
                            "Public Policy Offices") ~ 
               "Civil Society", 
             industry %in% c("Real Estate", 
                             "Facilities Services", 
                             "Environmental Services", 
                             "Recreational Facilities", 
                             "Leasing Non-residential Real Estate") ~ 
               "Real Estate", 
             industry == "Insurance" ~ "Insurance", 
             industry %in% c("Law Practice", 
                             "Legal Services", 
                             "Administration of Justice", 
                             "Alternative Dispute Resolution") ~ 
               "Legal", 
             industry %in%  c("Higher Education", 
                              "Education Administration Programs", 
                              "Primary and Secondary Education", 
                              "E-Learning Providers") ~ 
               "Education", 
             industry %in% c("Government Administration", 
                             "Legislative Offices", 
                             "Individual and Family Services") ~ 
               "Government Administration",
             industry %in% c("Biotechnology Research", 
                             "Pharmaceuticals Manufacturing", 
                             "Nanotechnology Research") ~ 
               "Pharmaceuticals and Biotech", 
             industry %in% c("Food and Beverage Services", 
                             "Restaurants", 
                             "Beverage Manufacturing", 
                             "Dairy Product Manufacturing") ~ 
               "F&B", 
             industry %in% c("Mental Health Care",  
                             "Hospitals and Health Care", 
                             "Medical Practices")
             ~ "Healthcare", 
             industry %in% c("Medical Equipment Manufacturing", 
                             "Industrial Machinery Manufacturing", 
                             "Appliances, Electrical, and Electronics Manufacturing", 
                             "Food and Beverage Manufacturing",
                             "Pharmaceutical Manufacturing", 
                             "Defense and Space Manufacturing", 
                             "Chemical Manufacturing", 
                             "Motor Vehicle Manufacturing", 
                             "Renewable Energy Semiconductor Manufacturing", 
                             "Aviation and Aerospace Component Manufacturing", 
                             "Automation Machinery Manufacturing", 
                             "Packaging and Containers Manufacturing", 
                             "Semiconductor Manufacturing", 
                             "Personal Care Product Manufacturing", 
                             "Computers and Electronics Manufacturing", 
                             "Plastics Manufacturing", 
                             "Furniture and Home Furnishings Manufacturing", 
                             "Paper and Forest Product Manufacturing", 
                             "Textile Manufacturing",
                             "Computer Hardware Manufacturing", 
                             "Sporting Goods Manufacturing", 
                             "Glass, Ceramics and Concrete Manufacturing", 
                             "Railroad Equipment Manufacturing", 
                             "Shipbuilding", 
                             "Tobacco Manufacturing", 
                             "Machinery Manufacturing", 
                             "Manufacturing")  ~ 
               "Manufacturing", 
             industry %in% c("Oil and Gas", 
                             "Mining") ~ 
               "Natural Resources", 
             industry %in% c("Utilities", 
                             "Telecommunications", 
                             "Wireless Services") ~ 
               "Telecoms and Utilities", 
             industry %in% c("Entertainment Providers", 
                             "Spectator Sports", 
                             "Broadcast Media Production and Distribution", 
                             "Media Production", 
                             "Book and Periodical Publishing",
                             "Gambling Facilities and Casinos", 
                             "Museums, Historical Sites, and Zoos", 
                             "Online Audio and Video Media", 
                             "Musicians", 
                             "Performing Arts", 
                             "Translation and Localization", 
                             "Newspaper Publishing", 
                             "Artists and Writers", 
                             "Libraries", 
                             "Photography",
                             "Animation and Post-production", 
                             "Writing and Editing") ~ 
               "Entertainment", 
             # Tell me it's not true
             industry %in% c("Airlines and Aviation", 
                             "Transportation, Logistics, Supply Chain and Storage", 
                             "Freight and Package Transportation", 
                             "Maritime Transportation", 
                             "Warehousing and Storage", 
                             "Truck Transportation") ~ 
               "Transportation and Logistics", 
             industry %in% c("Retail", 
                             "Retail Apparel and Fashion", 
                             "Consumer Services", 
                             "Retail Luxury Goods and Jewelry", 
                             # Maybe contentious, but Costco is a retailer
                             "Wholesale", 
                             "Retail Office Equipment",
                             "Retail Groceries", 
                             "Wholesale Import and Export") ~ 
               "Retail and Sales", 
             industry %in% c("Security and Investigations", 
                             "Public Safety", 
                             "Law Enforcement", 
                             "Armed Forces") ~ "Security", 
             industry %in% c("Travel Arrangements", 
                             "Hospitality") ~ 
               "Hospitality and Travel",
             industry %in% c("Veterinary Services", 
                             "Alternative Medicine", 
                             "Wellness and Fitness Services") ~ "Wellness"
           )) |> 
  # Just dropping all the extras
  group_by(company_id) |> 
  slice_head(n = 1) |> 
  ungroup()


job_category_cleaning <- function(tbl){
  tbl |>  
    mutate(
      job_category = 
        case_when(
          str_detect(title, "Chief|Director|Business Owner|Co-Founder|Executive|Vice President|Head") ~ 
                    "Director",
          str_detect(title, "Trainee|Intern|Paralegal|Entry Level|Support|Teller|ENTRY LEVEL|Accounting Clerk|Graduate Program|Universal Banker|Data Entry Clerk") ~ 
            "Entry Level",
          str_detect(title,
                     "Manager|MANAGER|Superintendent|Supervisor|Lead|Product Owner|Scrum Master|Cloud Architect|Sous Chef|Human Resources Business Partner|HR Business Partner")  & 
            !str_detect(title, "Chief") ~ 
            "Manager", 
          str_detect(title, "Engineer|Developer|Quality Inspector|Quality Control Inspector|Construction Inspector") ~ 
            "Engineer", 
          str_detect(title, 
                     "Attorney|Nurse|Therapist|Dental Hygienist|Architect|Licensed Clinical Social Worker|Physician|Nurse|NURSE|RN|Registered Dietitian|Chemist|Dentist|OCCUPATIONAL THERAPIST|Licensed Mental Health Counselor|Pharmacist|Psychiatrist|Veterinarian|Accountant|Bookkeeper|Auditor|Clinician|Behavior Interventionist|Neurologist|Cardiologist|Optometrist|THERAPIST|Medicine Resident") ~ 
                    "Licensed Professional",
          str_detect(title, "Consultant|Advisor") ~ 
                    "Consultant", 
          str_detect(title, "Technician|Detailer|Mechanic|Forklift Operator|Electrician|Machinist|Equipment Operator|Carpenter|Assembler|Operator|Medical Biller|Hair Stylist|Coordinator|Dispatcher| Tech|Plumber") ~ 
                    "Technician",
          str_detect(title, "Officer|Social Worker|Teacher|TEACHER|teacher|Caseworker") &
            !str_detect(title, "Chief") ~ 
                    "Officer",
          str_detect(title, 
                    "Specialist|Technologist|Controller|Phlebotomist|Scientist|Pathologist|Estimator|Data Architect|System Administrator|Systems Administrator|Database Administrator|Planner|Java   Architect|Network Administrator|Salesforce Administrator") ~ 
                    "Specialist", 
                  str_detect(title, "Analyst|Data Modeler") ~ 
            "Analyst",
          # Not sure about "Buyer"
                  str_detect(
                    title, "Associate|Salesperson|Team Member|Generalist|Merchandiser|Human Resources   Business Partner|Buyer|Relationship Banker|salesperson|Retail Parts Pro|ASSOCIATE|Sales Agent|Sales Professional|Sleep Expert - Sales|Representative|REPRESENTATIVE| Rep|Personal Trainer|PROMOTIONS AND SALES POSITIONS|Insurance Agent|Personal Banker|Recruiter|Store Counter Sales") ~
            "Associate",
                  str_detect(title, 
                             "Assistant|Receptionist|Retail Sales – Part Time|Front Desk Agent|ASSISTANT|Aide|Office Administrator|Contract Administrator|Legal Secretary|Branch Office Administrator|Payroll Administrator|Inventory Control Clerk|Accounts Payable Clerk") ~ 
            "Assistant",
                  str_detect(title, "Writer|writer|Content Editor|Designer|Copy Editor|Video Editor") ~ 
            "Creative",
          str_detect(title, "Cook|Driver|Housekeeper|Cashier|Dishwasher|Server|Bartender|Warehouse Worker|Janitor|Groomer|Food Service Worker|Welder|Custodian|Laborer|Dashers|Housekeeping   Aide|Barista|In-Store Shopper|Handler|Stocker|Drive with DoorDash|Caregiver|Warehouse Part Time Overnight|Cleaner|Warehouse Part Time Days|Restaurant Delivery|Warehouse Order Selector|Groundskeeper|Keyholder|Porter|Valet|store driver|Concierge|Room Attendant|Attendant|General Labor|Production Worker") ~ 
                    "Labourer", 
                  TRUE ~ "Other"
                  )) |> 
    mutate(senior = ifelse(str_detect(title, "Senior"), 
                           "senior", 
                           "not senior")) 
}

# path = kagglehub.dataset_download("arshkon/linkedin-job-postings")
# https://www.kaggle.com/datasets/arshkon/linkedin-job-postings


postings_clean <- read_csv("./raw_data/postings.csv") |>
  filter(!is.na(company_id)) |> 
  left_join(
    industries,
    by = "company_id"
  ) |> 
  filter(!is.na(industry)) |> 
  filter(!is.na(company_name) & company_name != "Confidential") |> 
  # Likely typos since there's entry-level and associates
  # Fine to drop them since it's less than 10 entries
  filter(normalized_salary < 1000000 | is.na(normalized_salary)) |> 
  job_category_cleaning()
```

I think let's ignore all staffing and recruiting postings. 



## By industry

We're looking at another set of scraped data today. This time, over 124,000 LinkedIn job postings over 2023-2024 were scraped by [Arsh Koneru and Zoey Yu Zou](https://www.kaggle.com/datasets/arshkon/linkedin-job-postings). 

Let's start 

```{r}
postings_clean |> 
  filter(industry_simple != "Recruiting") |> 
  group_by(industry_simple) |> 
  summarise(postings = n_distinct(job_id), 
            companies = n_distinct(company_id), 
            salary = mean(normalized_salary, na.rm = TRUE)) |> 
  ggplot(aes(x = postings, 
             y = fct_reorder(industry_simple, postings))) +
  geom_col(aes(fill = companies)) + 
  geom_text(aes(label = comma(salary)), 
            hjust = "inward") +
  scale_x_continuous(labels = comma) +
  scale_fill_viridis(direction = -1, option = "plasma", begin = .3) +
  labs(title = "LinkedIn job postings in 2023-2024, by industry", 
       subtitle = "Excludes postings by recruiters and middlemen. Mean salary (USD) at the end of each bar.", 
       y = "", x = "Number of job postings", 
       fill = "Number of\ncompanies", 
       caption = "Source: Arsh Koneru, Zoey Yu Zou and LinkedIn.com")
```

<br>

The IT and Healthcare industries have made the largest number of job postings on LinkedIn. IT and Manufacturing have the 

The IT industry continues to reward its workings comparatively well (to other industries, not to owners) and have more job postings than any other. Even with mass layoffs with the veneer of efficiency and AI, this trend probably continues to the present day. 

But I'm curious about the spread of jobs within each industry: which industries have the highest proportions of good jobs? And what even is a good job? 

<br><br><br>

## Job categories

Let's add a column for job categories to help determine which jobs are good jobs. 

Now, my categorisation is admittedly imperfect, given the large portion of jobs in the "Other" category. But there are diminishing returns to wading through that many layers of corporate obfuscation and doublespeak to determine where each job title should sit. 

But a cursory inspection of this categorisation, plotted below, shows that I have a fair grasp of job titles and what they actually mean. 



```{r}
postings_clean |> 
  filter(job_category == "Other") |> 
  count(title, sort = TRUE)


```



```{r}
postings_clean |>
  # Not sure this is the right filter
  # We can play around with it
  filter(normalized_salary > 1000 | is.na(normalized_salary)) |> 
  group_by(job_category) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE)) |> 
  ggplot(aes(x = n, 
             y = fct_reorder(job_category, n))) + 
  geom_col(aes(fill = salary)) +
  geom_text(aes(label = comma(salary)), 
            hjust = "inward") +
  scale_fill_viridis(direction = -1, labels = comma) + 
  scale_x_continuous(labels = comma) + 
  labs(x = "Number of postings", 
       y = "", 
       title = "Mean salary (USD) by job category",
       subtitle = "LinkedIn postings in 2023-2024", 
       fill = "Mean\nSalary", 
       caption = "Source: Arsh Koneru, Zoey Yu Zou and LinkedIn.com")

```

```{r}
postings_clean |>
  # Not sure this is the right filter
  # We can play around with it
  filter(normalized_salary > 1000 | is.na(normalized_salary)) |> 
  group_by(job_category) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE)) |> 
  arrange(desc(salary))
```

Plot by missing salary values or other values
What are we most blind to?
Money is the most important motivator. If it were so fun, I would do it for free. 


Below, we consider "good jobs" to be Directors, Engineers, Licensed Professionals and Managers, which are also the highest paying, on average.  

<br>

```{r}
postings_clean |> 
  filter(industry_simple != "Recruiting") |> 
  group_by(job_category, industry_simple) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE)) |> 
  group_by(industry_simple) |> 
  mutate(pc = n / sum(n),
         sort = sum(n[job_category %in% c(
           "Director", "Engineer", "Licensed Professional", "Manager")]) / 
           sum(n)) |> 
  ungroup() |> 
  mutate(job_category = fct_relevel(
    job_category, c("Director", "Engineer", "Licensed Professional", 
                    "Manager", "Consultant", "Creative", 
                    "Analyst", "Other", "Specialist", 
                    "Officer", "Associate", "Entry Level", 
                    "Technician", "Assistant", "Labourer")
  )) |> 
  arrange(desc(job_category)) |> 
  ggplot(aes(x = pc, y = fct_reorder(industry_simple, sort), 
             group = job_category)) + 
  geom_col(aes(fill = job_category)) + 
  scale_fill_viridis_d(option = "turbo") + 
  scale_x_continuous(labels = percent) +
  labs(title = "Breakdown of job postings by industry", 
       subtitle = "Sorted by percentage of 'good jobs'", 
       x = "", y = "", 
       fill = "") + 
  theme(legend.text = element_text(size = 9))
```

<br>

Retail and sales is one of the least top heavy industries. 

Education is dying (perhaps as a backlash towards its intensive monetisation and the purposeful destruction of the teaching profession), so is the legal industry, as well as government administration, which I can attest to have been in a state of atrophy even before DOGE and the great sack of the American government. These industries have an even worse proportion of good jobs than the oft-memed about Travel industry. 

Education is also still surprisingly top-heavy for a dying industry. 

We should probably cross-reference this with data scraped from Indeed, but the only available dataset is only a marketing sample of 30,000 jobs. 

Declining birthrates play a role in declining education. 

Unfortunately, this dataset is not scraped every year, so we can't monitor 

<br>

```{r}
category_plot <- function(tbl){
  
  tbl |> 
     mutate(job_category = fct_relevel(
      job_category, c("Director", "Engineer", "Licensed Professional", 
                      "Manager", "Consultant", "Creative", 
                      "Analyst", "Other", "Specialist", 
                      "Officer", "Associate", "Entry Level", 
                      "Technician", "Assistant", "Labourer")
    )) |>
    group_by(industry, job_category) |> 
    summarise(postings = n_distinct(job_id), 
              companies = n_distinct(company_id), 
              .groups = "drop") |> 
    ggplot(aes(x = postings, 
               y = fct_reorder(industry, postings), 
               group = job_category)) +
    geom_col(aes(fill = job_category)) + 
    scale_fill_viridis_d(option = "turbo")
}
```


```{r}
postings_clean |> 
  filter(industry_simple == "Recruiting") |> 
  group_by(company_name) |> 
  summarise(postings = n_distinct(job_id), 
            .groups = "drop") |>
  mutate(industry_postings = sum(postings), 
         pc = round(postings / industry_postings * 100, 2)) |> 
  arrange(desc(pc))
```

```{r}
postings_clean |> 
  filter(industry_simple == "Education") |> 
  category_plot()


```


<br>

Higher education is your last hope for a good job in the Education industry
Why there are more postings in Higher Education than Primary and Secondary Education speaks to? 
Supposedly, there is much more job security and less turnover in Primary and Secondary Education. 

Why are there, proportionally, so many more director-level jobs in Education?
Is Education too top-heavy? 
Massive monetisation of education. 

More likely, 
So maybe, it's more accurate to say that Higher Education (and Private Education) is dying


```{r}
postings_clean |> 
  filter(industry_simple == "Communications and Events") |> 
  group_by(industry, job_category) |> 
  summarise(n = n_distinct(job_id), 
            salary = mean(normalized_salary, na.rm = TRUE))
```





```{r}
postings_clean |>
  job_category_cleaning() |> 
  filter(job_category == "Manager") |> 
  count(title, sort = TRUE)
  filter(is.na(job_category)) |> 
  arrange(desc(n))
```

```{r}
postings_clean |> 
  job_category_cleaning() |> 
  filter(normalized_salary < 500000 ) |> 
  ggplot(aes(x = normalized_salary)) + 
  geom_histogram() 
  select(title, max_salary, med_salary, min_salary, pay_period, normalized_salary)
```

<br>

53% of LinkedIn users earn more than USD 100,000 a year. 54% of college graduates use LinkedIn. 

https://blog.hootsuite.com/linkedin-demographics/#:~:text=What%20the%20data%20tells%20us,makers%20and%20executive%20audiences%20worldwide.

## Industry consolidation and oligopolies

We live in trying times. The barrier to entry 

Don't put all your eggs in one basket

<br>

```{r}
postings_clean |> 
  filter(industry_simple != "Recruiting" & company_name != "A Hiring Company") |> 
  group_by(industry_simple, company_id) |> 
  summarise(postings = n_distinct(job_id), 
            .groups = "drop") |> 
  group_by(industry_simple) |> 
  mutate(industry_postings = sum(postings), 
         pc = postings / industry_postings) |> 
  filter(industry_postings > 100) |> 
  arrange(desc(pc)) |> 
  slice_head(n = 20) |>
  mutate(sort = sum(pc)) |> 
  ungroup() |> 
  ggplot(aes(x = pc, y = fct_reorder(industry_simple, sort))) + 
  geom_col(fill = "#adc178") + 
  scale_x_continuous(labels = percent, breaks = seq(0, .8, .1)) +
  labs(title = "What is the level of consolidation within industries?", 
       subtitle = "Industries ranked by the percentage of total jobs posted by the top 20 companies", 
       x = "", y = "")
    
  
```

<br>

```{r}
postings_clean |> 
  filter(industry_simple %in% c("Security", "Telecoms and Utilities", 
                                "Hospitality and Travel", "F&B", 
                                "Wellness")) |> 
  mutate(industry_simple = fct_relevel(industry_simple, 
                                       c("Security", "Telecoms and Utilities",
                                         "Hospitality and Travel", "F&B",
                                         "Wellness"))) |> 
  group_by(industry_simple, company_name) |> 
  summarise(postings = n_distinct(job_id), 
            .groups = "drop") |> 
  group_by(industry_simple) |> 
  mutate(industry_postings = sum(postings), 
         pc = round(postings / industry_postings * 100, 2)) |> 
  arrange(desc(pc)) |> 
  slice_head(n = 5) |>
  ungroup() |>
  select(-postings, -industry_postings) |> 
  kable(booktabs = TRUE, 
        caption = "Percent of postings by the five largest hirers per industry") |> 
  kable_classic_2("striped")
  
```

<br>

Whilst this is no [Herfindahl-Hirschman Index](https://en.wikipedia.org/wiki/Herfindahl%E2%80%93Hirschman_index#:~:text=HHI%20is%20calculated%20by%20squaring,share%2C%20weighted%20by%20market%20share.) (HHI), 

Though if you were to search up "Herfindahl-Hirschman Index by industry", you will find no relevant recent results. Perhaps of an indications that the oligopolistic tendencies of the corporate overclass have long gone unchecked and unquestioned. 

The last mention of it by the Bureau of Labor Statistics dates back to 2012 (with a report in 2014), and even then, the report never looks at consolidation and overconcentration of jobs within industries. But the position of the US government on oligopolies has long been clear. 

All industries show signs of oligopoly and excess concentration -- even in IT, 20 companies post 20% of all jobs. Notwithstanding the FAANG companies' stranglehold on patents and market capitalisation. 
 
So, what does this mean for you and me? If you were hoping that I would share some insights on how to pull the ladder up behind yourself,  

## Indeed

```{r}
indeed <- stream_in(file(("./data/marketing_sample_for_indeed_usa-indeed_usa_job__20210101_20210331__30k_data.ldjson")))

indeed |> glimpse()

indeed |> count(category, sort = TRUE)
```


